Model

### 2019-09-20 stats/model.Rmd
draft model comparison with noise parameter

@@

@@

@@

@@

  return {
    prag: pragModelPredictions.score(row.correct),
    globalPrag: globalPragModelPredictions.score(row.correct),
    flat: flatModelPredictions.score(row.correct),
    priorOnly: Math.log(row.correct == 1 ? row.prior : 1 - row.prior)
  }

@@

var log_likelihood_prag = sum(_.map(all_log_likes, "prag"))
var log_likelihood_prag_global = sum(_.map(all_log_likes, "globalPrag"))
var log_likelihood_flat = sum(_.map(all_log_likes, "flat"))
var log_likelihood_prior_only = sum(_.map(all_log_likes, "priorOnly"))

query.add(["log_like", "pragmatics", "NA"], ["NA", log_likelihood_prag])
query.add(["log_like", "pragmatics_global", "NA"], ["NA", log_likelihood_prag_global])
query.add(["log_like", "flat_prior", "NA"], ["NA", log_likelihood_flat])
query.add(["log_like", "prior_only", "NA"], ["NA", log_likelihood_prior_only])

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

saveRDS(sem_know_model, "../saves/sem_know_model_variant3_100k_model_comparison.rds")

@@

@@

```{r}
p_data_model <- ggplot(data = data_comb, aes(x = age_num, y = correct, col = source, fill = source, lty = alignment)) +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_jitter(width = 0, height = 0.05, alpha = .6)+
  #geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  geom_smooth(method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = comb_model_predictions, aes(x=age,y = me_effect, col = source, lty = alignment), size = 1)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(condition~familiarObject)+
  facet_grid(~item)+
  #ggtitle("Combination Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)

@@

```{r}
log_like<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "model","age","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -age, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(model,Chain)%>%
  summarize(logP = logSumExp(log_like))
```

@@

>>>>

var addNoise = function(dist, noiseParam){
   Infer({model: function(){ 
      return flip(noiseParam) ? uniformDraw([0, 1]) : sample(dist)
    }
   })
}

@@

var noise = uniformDrift({a: 0, b:1, width: 0.1})

@@

  var noisyPragModelPredictions = addNoise(pragModelPredictions, noise)

@@

   var noisyGlobalPragModelPredictions = addNoise(globalPragModelPredictions, noise)

@@

   var noisyFlatModelPredictions = addNoise(flatModelPredictions, noise)

  var priorOnlyModelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: prior}));
      return obj.shape == "novel_object" ? 1 : 0
       }})

  var noisyPriorOnlyModelPredictions = addNoise(priorOnlyModelPredictions, noise)

  return {
    prag: pragModelPredictions.score(row.correct),
    noisePrag: noisyPragModelPredictions.score(row.correct),

    globalPrag: globalPragModelPredictions.score(row.correct),
    noiseGlobalPrag: noisyGlobalPragModelPredictions.score(row.correct),

    flat: flatModelPredictions.score(row.correct),
    noiseFlat: noisyFlatModelPredictions.score(row.correct),

    priorOnly: priorOnlyModelPredictions.score(row.correct),
    noisePriorOnly: noisyPriorOnlyModelPredictions.score(row.correct)
  }

@@

var log_likelihood_prag = sum(_.map(all_log_likes, "prag"))
var log_likelihood_prag_noise = sum(_.map(all_log_likes, "noisePrag"))

var log_likelihood_prag_global = sum(_.map(all_log_likes, "globalPrag"))
var log_likelihood_prag_global_noise = sum(_.map(all_log_likes, "noiseGlobalPrag"))

var log_likelihood_flat = sum(_.map(all_log_likes, "flat"))
var log_likelihood_flat_noise = sum(_.map(all_log_likes, "noiseFlat"))

var log_likelihood_prior_only = sum(_.map(all_log_likes, "priorOnly"))
var log_likelihood_prior_only_noise = sum(_.map(all_log_likes, "noisePriorOnly"))

query.add(["log_like", "pragmatics", "parameter_free"], ["NA", log_likelihood_prag])
query.add(["log_like", "pragmatics", "noise"], ["NA", log_likelihood_prag_noise])

query.add(["log_like", "pragmatics_global", "parameter_free"], ["NA", log_likelihood_prag_global])
query.add(["log_like", "pragmatics_global", "noise"], ["NA", log_likelihood_prag_global_noise])

query.add(["log_like", "flat_prior", "parameter_free"], ["NA", log_likelihood_flat])
query.add(["log_like", "flat_prior", "noise"], ["NA", log_likelihood_flat_noise])

query.add(["log_like", "prior_only", "parameter_free"], ["NA", log_likelihood_prior_only])
query.add(["log_like", "prior_only", "noise"], ["NA", log_likelihood_prior_only_noise])

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 25000,
                        burn = 20000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

saveRDS(sem_know_model, "../saves/sem_know_model_variant3_25k_model_comparison_noise.rds")

@@

```{r}
log_like<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "model","noise","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(model,noise)%>%
  summarize(logP = logSumExp(log_like))
```

@@

```{r}
p_data_model <- ggplot(data = data_comb, aes(x = age_num, y = correct, col = alignment, fill = alignment, lty = source)) +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_jitter(width = 0, height = 0.05, alpha = .6)+
  #geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  geom_smooth(method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = comb_model_predictions, aes(x=age,y = me_effect, col = alignment, lty = source), size = 1)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(condition~familiarObject)+
  facet_grid(~item)+
  #ggtitle("Combination Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)

@@

@@

# MAP based predictions with noise parameter

## Preparing data

```{r}
noise_model_data <- item_params_summary%>%
  select(-uci,-lci)%>%
  spread(parameter, mode) %>%
  left_join(read_csv("../data/combination.csv") %>%
  filter(condition != "train")%>%
  mutate(alignment = condition,
         item = familiarObject,
         age = age_num - (min(age_num))))%>%
  mutate(so_int = global_params_summary%>%filter(token == "speaker_optimality", parameter == "int")%>%pull(mode),
         so_slope = global_params_summary%>%filter(token == "speaker_optimality", parameter == "slope")%>%pull(mode),
    glob_sem_know_int = global_params_summary%>%filter(token == "global_sem", parameter == "int")%>%pull(mode),
         glob_sem_know_slope = global_params_summary%>%filter(token == "global_sem", parameter == "slope")%>%pull(mode),
         speaker_optimality = so_int + so_slope * age,
         semantic_knowledge = plogis(int + slope * age),
         global_semantic_knowledge = plogis(glob_sem_know_int + glob_sem_know_slope * age),
         nov_int = novel_params$Intercept,
         nov_slope = novel_params$age_num,
         prior = ifelse(alignment == "congruent", plogis(nov_int + nov_slope* age),1-plogis(nov_int + age * nov_slope)))%>%
  select(item, alignment,age, speaker_optimality, semantic_knowledge,global_semantic_knowledge, prior, correct)

## Pragmatics model

```{r}
pragNoise <- '
var allData = dataFromR.data

  var noise = uniform({a: 0, b: 1})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct)

    return extend([row.alignment + "/" +  row.item + "/" + row.age, Math.exp(noisyModelPredictions.score(1))])

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r predictions pragmatic model ex 3 adults}
prag_model_noise<- webppl(
  program_code = paste(rsaUtils, rsaModel, pragNoise , sep='\n'),
  data =list(data = noise_model_data),
  data_var = "dataFromR",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 1000, verbose = T)
)

saveRDS(prag_model_noise, "../saves/prag_model_noise.rds")

prag_model_noise <- readRDS("../saves/prag_model_noise.rds")

# noise parameter
prag_model_noise_noise_param_summary <- prag_model_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

# model predictions
prag_model_noise_predictions <- prag_model_noise %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("alignment","item", "age"), sep="/")%>%
  mutate(model = "pragmatic_noise")%>%
  group_by(model,age,item, alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         source = "Model")

### Plot model predictions

```{r}
ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb, aes(x = age_num, y = correct, col = alignment, fill = alignment, lty = source), method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = prag_model_noise_predictions, aes(x=age+2,y = mean, col = alignment, lty = source), size = 1)+
  geom_ribbon(data = prag_model_noise_predictions, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_wrap(~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)

## Pragmatics model global semantic knnowledge

```{r}
globalPragNoise <- '
var allData = dataFromR.data

  var noise = uniform({a: 0, b: 1})

    var modelPredictions = pragmaticListener({label: "novel_word"}, prior, row.speaker_optimality,row.global_semantic_knowledge)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct)

    return extend([row.alignment + "/" +  row.item + "/" + row.age, Math.exp(noisyModelPredictions.score(1))])

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r predictions pragmatic model ex 3 adults}
global_prag_model_noise<- webppl(
  program_code = paste(rsaUtils, rsaModel, globalPragNoise , sep='\n'),
  data =list(data = noise_model_data),
  data_var = "dataFromR",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 1000, verbose = T)
)

saveRDS(global_prag_model_noise, "../saves/global_prag_model_noise.rds")

global_prag_model_noise <- readRDS("../saves/global_prag_model_noise.rds")

# noise parameter
global_prag_model_noise_noise_param_summary <- global_prag_model_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

# model predictions
global_prag_model_noise_predictions <- global_prag_model_noise %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("alignment","item", "age"), sep="/")%>%
  mutate(model = "global_pragmatic_noise")%>%
  group_by(model,age,item, alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         source = "Model")

### Plot model predictions

```{r}
ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb, aes(x = age_num, y = correct, col = alignment, fill = alignment, lty = source), method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = global_prag_model_noise_predictions, aes(x=age+2,y = mean, col = alignment, lty = source), size = 1)+
  geom_ribbon(data = global_prag_model_noise_predictions, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_wrap(~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)

## Flat prior model

```{r}
flatNoise <- '
var allData = dataFromR.data

  var noise = uniform({a: 0, b: 1})

    var modelPredictions = pragmaticListener({label: "novel_word"}, [.5,.5], row.speaker_optimality,row.semantic_knowledge)

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct)

    return extend([row.alignment + "/" +  row.item + "/" + row.age, Math.exp(noisyModelPredictions.score(1))])

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r predictions pragmatic model ex 3 adults}
flat_model_noise<- webppl(
  program_code = paste(rsaUtils, rsaModel, flatNoise , sep='\n'),
  data =list(data = noise_model_data),
  data_var = "dataFromR",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 1000, verbose = T)
)

saveRDS(flat_model_noise, "../saves/flat_model_noise.rds")

flat_model_noise <- readRDS("../saves/flat_model_noise.rds")

# noise parameter
flat_model_noise_noise_param_summary <- flat_model_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

# model predictions
flat_model_noise_predictions <- flat_model_noise %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("alignment","item", "age"), sep="/")%>%
  mutate(model = "flat_prior_noise")%>%
  group_by(model,age,item, alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         source = "Model")

### Plot model predictions

```{r}
ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb, aes(x = age_num, y = correct, col = alignment, fill = alignment, lty = source), method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = flat_model_noise_predictions, aes(x=age+2,y = mean, col = alignment, lty = source), size = 1)+
  geom_ribbon(data = flat_model_noise_predictions, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_wrap(~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```

## Prior only model

```{r}
priorOnlyNoise <- '
var allData = dataFromR.data

  var noise = uniform({a: 0, b: 1})

      var modelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: prior}));
      return obj.shape == "novel_object" ? 1 : 0
       }})

    var noisyModelPredictions = addNoise(modelPredictions, noise)

    observe(noisyModelPredictions, row.correct)

    return extend([row.alignment + "/" +  row.item + "/" + row.age, Math.exp(noisyModelPredictions.score(1))])

 return extend(_.fromPairs(output), {noise: noise})

}
'
```

```{r predictions pragmatic model ex 3 adults}
prior_only_model_noise<- webppl(
  program_code = paste(rsaUtils, rsaModel, priorOnlyNoise , sep='\n'),
  data =list(data = noise_model_data),
  data_var = "dataFromR",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "MCMC", samples = 2000, burn = 1000, verbose = T)
)

saveRDS(prior_only_model_noise, "../saves/prior_only_model_noise.rds")

prior_only_model_noise <- readRDS("../saves/prior_only_model_noise.rds")

# noise parameter
prior_only_model_noise_noise_param_summary <- prior_only_model_noise %>%
  filter(Parameter %in% c("noise"))  %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value),
            ci_upper = hdi_upper(value))

# model predictions
prior_only_model_noise_predictions <- prior_only_model_noise %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("alignment","item", "age"), sep="/")%>%
  mutate(model = "prior_only_noise")%>%
  group_by(model,age,item, alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         source = "Model")

### Plot model predictions

```{r}
ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb, aes(x = age_num, y = correct, col = alignment, fill = alignment, lty = source), method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = prior_only_model_noise_predictions, aes(x=age+2,y = mean, col = alignment, lty = source), size = 1)+
  geom_ribbon(data = prior_only_model_noise_predictions, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_wrap(~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```

## Compare noise parameters

```{r}
noise_plot <- bind_rows(
  prag_model_noise %>%
    filter(Parameter %in% c("noise"))%>%
    mutate(Model = "Pragmatic"),
  global_prag_model_noise %>%
    filter(Parameter %in% c("noise"))%>%
    mutate(Model = "Pragmatic no word knowledge"),
  flat_model_noise %>%
    filter(Parameter %in% c("noise"))%>%
    mutate(Model = "Flat Prior"),
  prior_only_model_noise %>%
    filter(Parameter %in% c("noise"))%>%
    mutate(Model = "Prior Only")
)

ggplot(noise_plot, aes(x = value, col = Model, fill = Model))+
  geom_density(alpha = .7)+
  theme_few()+
  xlim(0,1)+
  scale_color_ptol()+
  scale_fill_ptol()+
  theme(legend.position = "bottom")

## Compare predictions

noise_model_pred <- bind_rows(
  prag_model_noise_predictions,
  global_prag_model_noise_predictions,
  prior_only_model_noise_predictions,
  flat_model_noise_predictions
)

data_comb_plot<-data_comb%>%
  mutate(model = "data")

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct, col = "black", fill = model), method = "loess", se = F, alpha = .2, span = 2)+
  geom_line(data = noise_model_pred, aes(x=age+2,y = mean, col = model), size = 1)+
  geom_ribbon(data = noise_model_pred, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```
